Testing Set Features Visualization

### 2022-11-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations for SI.

```{r}
ggplot(sm_m, aes(x=method, y=score,
                   fill=score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Viral Score',
                     values = alpha(c(viridis(6)), 1)) +
  xlab("") +
  ylab("Number of Sequences") +
  coord_flip()
```
## Visualizing confusion matrix by number of tools

@@

>>>>

```{r}
ggplot(sm_m, aes(x=method, y=score,
                   fill=score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Viral Score',
                     values = alpha(c(viridis(6)), 1)) +
  xlab("") +
  ylab("Number of Sequences") +
  coord_flip()
```
## Another way of visualizing between rule sets

```{r}
viruses_mcc_alluvial <- data.frame(seqtype=viruses$seqtype,
                                   keep_score_high_MCC=viruses$keep_score_high_MCC,
                                   confusion_matrix_high_MCC=viruses$confusion_matrix_high_MCC)

viruses_mcc_alluvial$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_mcc_alluvial$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_mcc_alluvial$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses_mcc_alluvial$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_mcc_alluvial$keep_score_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_mcc_alluvial$keep_score_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
```

```{r}
viruses_mcc_alluvial %>%
  count(seqtype, keep_score_high_MCC) %>% 
  spread(key = keep_score_high_MCC, value=n)
```

```{r}
viruses_mcc_alluvial <- viruses_mcc_alluvial %>%
  count(seqtype, keep_score_dvf, keep_score_vb, keep_score_vs,
        keep_score_vs2, keep_score_tv, keep_score_tnv, keep_score_high_MCC) %>%
  mutate(high_mcc_viral_score=factor(round(keep_score_high_MCC)))
```

```{r}
ggplot(viruses_mcc_alluvial,
       aes(axis1 = keep_score_dvf, axis2 = keep_score_vb, 
           axis3 = keep_score_vs, axis4 = keep_score_vs2, 
           axis5 = keep_score_tv, axis6 = keep_score_tnv, 
           y=n)) +
  geom_alluvium(aes(fill=high_mcc_viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4,5,6),
    labels=c("dvf", "kj", "vs", "vs2",
             "tv", "tnv")) +
  scale_fill_brewer(palette = "PuOr", ) +
  facet_wrap(~seqtype, scales="free_y") 
```

## Visualizing confusion matrix by number of tools

@@

```{r}
confusion_by_taxa <- viruses %>% count(confusion_matrix_high_recall, seqtype, size_class)

colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "count")
```

### 2022-11-07 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
New visualizations for SI.

use of each tuning rule
```{r}
keep_score <- rep(0, nrow(viruses))

keep_score[viruses$hallmark>2] <- keep_score[viruses$hallmark>2] + 1
table(keep_score, viruses$seqtype)
keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] - 1
table(keep_score, viruses$seqtype)
keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
table(keep_score, viruses$seqtype)
keep_score[viruses$percent_viral>=50] <- keep_score[viruses$percent_viral>=50] + 0.5
table(keep_score, viruses$seqtype)
    #keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] <- keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] + 1 #add some ratio
keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
table(keep_score, viruses$seqtype)
keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] - 1 # consider accounting for provirus designation
table(keep_score, viruses$seqtype)
#    keep_score[(viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes] <- 0 # consider accounting for provirus designation
keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] <- keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] - 1
table(keep_score, viruses$seqtype)
```

>>>>

# Viral Addition Tuning Rules

```{r}
viruses_sankey_tv <- data.frame(seqtype=viruses$seqtype,
                             kj_cel=rep(0,nrow(viruses)),
                             hall=rep(0,nrow(viruses)),
                             pv=rep(0,nrow(viruses)),
                             cvl_pu=rep(0, nrow(viruses)))
```

```{r}
viruses_sankey_tv$kj_cel[viruses$Kaiju_Viral=="Viruses"] <- 0.5
viruses_sankey_tv$hall[viruses$hallmark>2] <- 0.5
viruses_sankey_tv$pv[viruses$percent_viral>=50] <- 0.5
viruses_sankey_tv$cvl_pu[viruses$checkv_length>50000 & viruses$percent_unknown<=75] <- 0.5

viruses_sankey_tv$all <- rowSums(viruses_sankey_tv[,2:5])
```

```{r}
viruses_sankey_tv %>%
  count(seqtype, all) %>% spread(key = all, value=n)
```

```{r}
viruses_sankey_tv <- viruses_sankey_tv %>%
  count(seqtype, kj_cel, hall, pv, cvl_pu, all) %>%
  mutate(viral_score=factor(all))
```

```{r}
ggplot(viruses_sankey_tv,
       aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
  geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4),
    labels=c("kaiju", "hallmark", "% viral", "% unknown")) +
  facet_wrap(~seqtype, scales="free_y") 
```

# Viral Removal Tuning Rules

```{r}
viruses_sankey_tnv <- data.frame(seqtype=viruses$seqtype,
                             kj_cel=rep(0,nrow(viruses)),
                             hg_pro=rep(0,nrow(viruses)),
                             vg_hg=rep(0,nrow(viruses)),
                             vg_hg_pro=rep(0, nrow(viruses)),
                             cvl_hm=rep(0, nrow(viruses)),
                             cvl_cp=rep(0,nrow(viruses)))
```

```{r}
viruses_sankey_tnv$kj_cel[viruses$Kaiju_Viral=="cellular organisms"] <- -0.5
viruses_sankey_tnv$hg_pro[viruses$checkv_host_genes>50 & viruses$provirus==FALSE] <- -1
viruses_sankey_tnv$vg_hg[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- -1
viruses_sankey_tnv$vg_hg_pro[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus==FALSE] <- -1
viruses_sankey_tnv$cvl_hm[viruses$checkv_length>500000 & viruses$hallmark<=1] <- -1
viruses_sankey_tnv$cvl_cp[viruses$checkv_length>5000 & viruses$checkv_completeness<=75] <- -0.5             
viruses_sankey_tnv$all <- rowSums(viruses_sankey_tnv[,2:7])
```

```{r}
viruses_sankey_tnv %>%
  count(seqtype, all) %>% spread(key = all, value=n)
```

```{r}
viruses_sankey_tnv <- viruses_sankey_tnv %>%
  count(seqtype, kj_cel, hg_pro, vg_hg, vg_hg_pro, cvl_hm, cvl_cp, all) %>%
  mutate(viral_score=factor(all))
```

```{r}
ggplot(viruses_sankey_tnv,
       aes(axis1 = kj_cel, axis2 = hg_pro, axis3 = vg_hg, axis4 = cvl_hm, axis5=cvl_hm, axis6=cvl_cp, 
           y=n)) +
  geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4,5,6),
    labels=c("kaiju", "host genes", "viral and host", "provirus",
             "hallmark", "complete")) +
  facet_wrap(~seqtype, scales="free_y") 
```
